### 2025-05-07 chapters/2-particles.tex
Completed problems and rough solutions

In Yang-Mills theory with gauge group SU(N), we instead write \(U_{(R)} = e^{-i g T_{(R)}^a \alpha^a(x)}\) is upgraded to an element of (a representation of) SU(N), where the \(\{T^a_{(R)}\}\) are linearly-independent, traceless matrices:
%
the generators of SU(N) in the representation \(R\).
%
The index \(a\) takes values in \(a \in \{1,\,\cdots,\,N^2-1\}\), and in any representation the generators are guaranteed to satisfy the condition
\begin{align}
    \comm{T^a_{(R)}}{T^b_{(R)}}
    =
    i f^{abc} T^c_{(R)}
    \,,
\end{align}
for geometric reasons that we do not cover here.
%
The \(f^{abc}\) are numbers, and are called the \vocab{structure constants} of the group.
%
The generators are conventionally chosen to satisfy
\begin{align}
    \Tr\le(T_{(R)}^a T_{(R)}^b\ri) = T_R \delta^{ab}
    \,,
\end{align}
where the factor 1/2 is called the \textit{Dynkin index} (after mathematician Eugene Dynkin), which we will discuss more when we introduce the Feynman rules of \gls{qcd}.
%
These requirements on the generators can be used to show that the \(f^{abc}\) are completely antisymmetric with respect to exchange of any two indices.

@@

Particularly notable for the case of \gls{qcd} is the beta function for the Yang-Mills coupling \(g_s\), called the \textit{strong coupling constant}.
%
It is often parameterized in terms of perturbatively calculable coefficients as
\begin{align}
    \beta_{g_s}(g_s)
    =
    -g_s\,
    \sum_{n=0}^\infty \beta_n \le(\frac{g_s^2}{16\pi^2}\ri)^{n+1}
    \,,
\end{align}
or
\begin{align}
    \beta_{\alpha_s}(\alpha_s)
    =
    -2\alpha_s
    \sum_{n=0}^\infty \beta_n \le(\frac{\alpha_s}{4\pi}\ri)^{n+1}
\end{align}
where \(\alpha_s = g_s^2 / 4 \pi\) and the \(\beta_k\) are coefficients that can be captured by N\(^{k+1}\)LO computations.
%
This parameterization is used because the combination \(\alpha_s / 4\pi\) appears frequently in loop calculations.

@@

    Show that the phase space factor for a single final-state particle can be written in the manifestly (orthocronous) Lorentz invariant form
    \begin{align}
        \frac{\dd^{d-1} k}{\le(2\pi\ri)^{d-1} 2 E_k}
        =
        \int
        \frac{\dd^d k}{\le(2\pi\ri)^d}
        \Theta(k^0 > 0) \delta\le(k^2 - m^2\ri)
        \,,
    \end{align}
    where the integral of the right-hand-side is done over \(k^0\) (and is often left out in writing).

@@

    \begin{subequations}
    \begin{align}
        \label{eq:p_gtogg}
        P_{gg \leftarrow gg}(z)
        &=
        2C_A \le(
            \frac{1-z}{(z)_+}
            +
            \frac{z}{(1-z)_+}
            +
            z(1-z)
        \ri)
        \,\,
        +
        \,\,
        \frac{\beta_0}{2}
        \,
        \le(\delta(z) + \delta(1-z)\ri)
        \,,
        \\
        \label{eq:p_qtogq}
        P_{gq \leftarrow q}(z)
        &=
        C_F \le[\frac{1+\le(1-z\ri)^2}{z}\ri]_+
        \,,
        \\
        \label{eq:p_gtoqq}
        P_{q\bar{q} \leftarrow g}(z)
        &=
        T_F \left(z^2 + (1-z)^2\right)
        \,.
    \end{align}
    \end{subequations}
\end{answer}

@@

\begin{lemma}{Angular Ordering}{angular-ordering}
    The integral of the antenna factor \(A_{ij}\) over the solid-angle direction of \(\vec{q}\) can be expressed as the sum of two pieces,
    \begin{align}
        \label{eq:angular-ordering-integral}
        \int \frac{\dd \Omega}{2\pi}\, A_{ij}
        &=
        \int_{\cos\theta_{ij}}^{1}
        \frac{\dd \cos\theta_{iq}}{1-\cos\theta_{iq}}
        +
        \int_{\cos\theta_{ij}}^{1}
        \frac{\dd \cos\theta_{jq}}{1-\cos\theta_{jq}}
        \\
        &=
        2
        \le(
            \int_{0}^{\theta_{ij}}
            \frac{\dd \theta_{iq}}{\theta_{iq}}
            +
            \int_{0}^{\theta_{ij}}
            \frac{\dd \theta_{jq}}{\theta_{jq}}
        \ri)
        +
        \mathcal{O}(\theta_{ij}^2)
        \,,
    \end{align}
    where the first term of the right-hand-side emerges after integrating over the azimuthal direction of \(\vec{q}\) relative to the axis defined by the motion of particle \(i\), and the second when integrating over the azimuthal angle relative to \(j\).
    %
    Notably, the phase space measure for \(\vec{q}\) of which vanishes when the direction of \(\vec{q}\) lies outside the cone defined by the momenta of \(i\) (in the first term) or \(j\) (in the second).
\end{lemma}

@@

    In the limit that \(q\) is collinear to particle 1, is it the first term in the parentheses of the right-hand-side that matters.

@@

\begin{lemma}{Coherent Branching}{coherent-branching}
    Consider \(M\) outgoing particles, and an additional soft gluon, in the \vocab{strongly-angular-ordered limit}:%
    \footnote{
        Using the triangle equality, we could also write
        \(
            \theta_{12} \ll \theta_{(12)3} \ll \theta_{(123)4} \ll \cdots
            \,,
        \)
        where \(\theta_{(S)\,\,n}\) is the angle between \(\sum_{i\in S} \vec{k}_i\) and \(\vec{k}_n\).
    }
    \begin{align}
        \theta_{2} \ll \theta_{3} \ll \theta_{4}
        \ll
        \cdots
        \ll
        \theta_M
        \,,
    \end{align}
    where we define \(\theta_j = \theta_{1j}\) to be the angle of particle \(j\) relative to particle 1 for brevity.

@@

%\subsection{Splitting Functions}

%Finally, we tabulate the properties of the exclusive and inclusive \gls{qcd} \glspl{splittingfn} in \Tabs{exclusive-splitting}{inclusive-splitting} respectively.
%%
%We use \(B(m,n)\) to denote Euler's beta function,
%\begin{align}
%    B(m,n) = \frac{\Gamma(m)\Gamma(n)}{\Gamma(m+n)}
%    \,.
%\end{align}

%\sam{Missing delta functions/virtual contributions}
%%

%% +===================================
%\begin{table}[h!]
%\begin{center}
%\LARGE $\boldsymbol{P_{ij \leftarrow \ell}(z)}$
%\end{center}
%\vspace{-10pt}
%\centering
%\caption[Exclusive NLO QCD splitting functions and their Mellin moments.]{
%    Exclusive QCD splitting functions and their Mellin moments at NLO (\(z = E_i/E_\ell\)).
%    %
%    $B(m,n)$ denotes Euler's beta function.  \sam{check}
%    %
%    \sam{include comment about \(N_f\)}
%}
%\label{tab:exclusive-splitting}
%\vspace{5pt}
%\renewcommand{\arraystretch}{3.2}
%\rowcolors{1}{gray!20}{gray!5}
%\begin{tabular}{|>{\bfseries}m{2cm}|m{4cm}|m{7cm}|}
%\hline
%\centering \textbf{Channel} & \centering \textbf{Splitting Function} &
%\centering $\boldsymbol{\hat{P}(m,n)} := \int_0^1 \dd x
%\, x^{m-1} \le(1-x\ri)^{n-1} \, \, P(x)$
%\tabularnewline
%\hline
%\centering $gq \leftarrow q$
%                            &
%\centering $C_F \dfrac{1 + (1 - z)^2}{z}$
%                            &
%\centering
%$C_F \le(B\le(m,n-1\ri) + B\le(m+2,n-1\ri)\ri)$
%\tabularnewline
%\hline
%\centering
%$gg \leftarrow g$
%                            &
%\centering $2 C_A \dfrac{(1 - z(1 - z))^2}{z(1 - z)}$
%                            &
%\centering
%$2 C_A \big(B\le(m-1, n+1\ri) + B\le(m+1, n-1\ri)$
%\centering
%$+ B\le(m+1, n+1\ri)\big)$
%\tabularnewline
%\hline
%\centering $q\bar{q} \leftarrow g$
%                            &
%\centering $T_F \le(z^2 + (1-z)^2\ri)$
%                            &
%\centering
%$T_F \le(B\le(m+2, n\ri) + B\le(m, n+2\ri)\ri)$
%\tabularnewline
%\hline
%\end{tabular}
%\end{table}
%% +===================================

%\vspace{1em}
%% +===================================
%\begin{table}[h!]
%\begin{center}
%\LARGE $\boldsymbol{p_{i \leftarrow \ell}(z)}$
%\end{center}
%\vspace{-10pt}
%\centering
%\caption[Inclusive NLO QCD splitting functions and their Mellin moments.]{
%    Inclusive QCD Splitting Functions and their Mellin moments at NLO.  \sam{check}
%    %
%    \sam{comment about \(N_f\)}
%}
%\label{tab:inclusive-splitting}
%\vspace{5pt}
%\renewcommand{\arraystretch}{3.5}
%\rowcolors{1}{gray!20}{gray!5}
%\begin{tabular}{|>{\bfseries}m{1.5cm}|m{6cm}|m{6cm}|}
%\hline
%\centering \textbf{Channel} & \centering \textbf{Splitting Function} & \centering $\boldsymbol{\hat{p}(m)} := \int \dd x\, x^{m-1} \, \, p(m)$
%\tabularnewline
%\hline
%\centering $q \leftarrow q$
%                            &
%\centering $C_F \left[ \dfrac{1 + z^2}{1 - z} \right]_+$
%                            &
%\centering $C_F \left(-\dfrac{1}{2} + \dfrac{1}{m(m+1)} - 2\sum_{k=2}^{m} \dfrac{1}{k} \right)$
%\tabularnewline
%\hline
%\centering $g \leftarrow q$
%                            &
%\centering $C_F \left( \dfrac{1 + (1 - z)^2}{z} \right)$
%                            &
%\centering
%$C_F \dfrac{m^2 + m + 2}{m(m-1)(m+1)}$
%\tabularnewline
%\hline
%\centering $g \leftarrow g$
%                            &
%\centering $2C_A \left[ \dfrac{z}{(1 - z)_+} + \dfrac{1 - z}{z} + z(1 - z) \right]$
%                            &
%\centering
%$C_A\le(-\frac{1}{6} + \frac{2}{n(n+1)} + \frac{2}{(n+2)(n+3)}\ri) - 2C_A \sum_{k=2}^{m} \dfrac{1}{k} - N_f / 3$
%\tabularnewline
%\hline
%\centering
%$q \leftarrow g$
%($\overline{q} \leftarrow g$)
%                            &
%\centering $T_F \left( z^2 + (1 - z)^2 \right)$
%                            &
%\centering
%$2T_F \dfrac{m^2+3m+4}{(m+1)(m+2)(m+3)}$
%\tabularnewline
%\hline
%\end{tabular}
%\end{table}
%% +===================================

@@

\makeprob{The Cauchy Principle Value}{principle-value}{
    Provide a formal argument that
    \begin{align}
        \frac{\dd}{\dd x} \ln\le(\abs{x}\ri)
        =
        \text{PV}\le(\frac{1}{x} \ri)
        \,,
    \end{align}
    where the right-hand-side is the \vocab{Cauchy Principle Value}, a distribution defined by
    \begin{align}
        \int_{-\infty}^{\infty} \dd x
        \,\,
        \text{PV}\le(\frac{1}{x} \ri)
        \,
        \phi(x)
        =
        \int_0^\infty
        \,
        \frac{\phi(x) - \phi(-x)}{x}
        \,
        \dd x
        \,.
    \end{align}
}

>>>>

In Yang-Mills theory with gauge group SU(N), we instead write \(U_{(R)} = e^{-i g T_{(R)}^a \alpha^a(x)}\) is upgraded to an element of (a representation of) SU(N), where the \(\{T^a_{(R)}\}\) are linearly-independent, traceless matrices:
%
the generators of SU(N) in the representation \(R\).
%
The index \(a\) takes values in \(a \in \{1,\,\cdots,\,N^2-1\}\), and in any representation the generators are guaranteed to satisfy the condition
\begin{align}
    \comm{T^a_{(R)}}{T^b_{(R)}}
    =
    i f^{abc} T^c_{(R)}
    \,,
\end{align}
for geometric reasons that we do not cover here.
%
The \(f^{abc}\) are numbers, and are called the \vocab{structure constants} of the group.
%
The generators are conventionally chosen to satisfy
\begin{align}
    \Tr\le(T_{(R)}^a T_{(R)}^b\ri) = T_R \delta^{ab}
    \,,
\end{align}
where the factor \(T_R\) is called the \textit{Dynkin index} (after mathematician Eugene Dynkin), which we will discuss more when we introduce the Feynman rules of \gls{qcd}.
%
These requirements on the generators can be used to show that the \(f^{abc}\) are completely antisymmetric with respect to exchange of any two indices.

@@

Particularly notable for the case of \gls{qcd} is the beta function for the Yang-Mills coupling \(g_s\), called the \textit{strong coupling constant}.
%
It is often parametrized in terms of perturbatively calculable coefficients as
\begin{align}
    \beta_{g_s}(g_s)
    =
    -g_s\,
    \sum_{n=0}^\infty \beta_n \le(\frac{g_s^2}{16\pi^2}\ri)^{n+1}
    \,,
\end{align}
or
\begin{align}
    \beta_{\alpha_s}(\alpha_s)
    =
    -2\alpha_s
    \sum_{n=0}^\infty \beta_n \le(\frac{\alpha_s}{4\pi}\ri)^{n+1}
\end{align}
where \(\alpha_s = g_s^2 / 4 \pi\) and the \(\beta_k\) are coefficients that can be captured by N\(^{k+1}\)LO computations.
%
This parametrization is used because the combination \(\alpha_s / 4\pi\) appears frequently in loop calculations.

@@

    Show that the phase space factor for a single final-state particle can be written in the manifestly (orthocronous) Lorentz invariant form
    \begin{align}
        \frac{\dd^{d-1} k}{\le(2\pi\ri)^{d-1} 2 E_k}
        =
        \int
        \frac{\dd^d k}{\le(2\pi\ri)^d}
        \Theta(k^0 > 0) \delta\le(k^2 - m^2\ri)
        \,,
    \end{align}
    where the integral of the right-hand side is done over \(k^0\) (and is often left out in writing).

@@

    \begin{subequations}
    \begin{align}
        \label{eq:p_gtogg}
        P_{gg \leftarrow g}(z)
        &=
        2C_A \le(
            \frac{1-z}{(z)_+}
            +
            \frac{z}{(1-z)_+}
            +
            z(1-z)
        \ri)
        \,\,
        +
        \,\,
        \frac{\beta_0}{2}
        \,
        \le(\delta(z) + \delta(1-z)\ri)
        \,,
        \\
        \label{eq:p_qtogq}
        P_{gq \leftarrow q}(z)
        &=
        C_F \le[\frac{1+\le(1-z\ri)^2}{z}\ri]_+
        \,,
        \\
        \label{eq:p_gtoqq}
        P_{q\bar{q} \leftarrow g}(z)
        &=
        T_F \left(z^2 + (1-z)^2\right)
        \,.
    \end{align}
    \end{subequations}
\end{answer}

@@

\begin{lemma}{Angular Ordering}{angular-ordering}
    The integral of the antenna factor \(A_{ij}\) over the solid-angle direction of \(\vec{q}\) can be expressed as the sum of two pieces,
    \begin{align}
        \label{eq:angular-ordering-integral}
        \int \frac{\dd \Omega}{2\pi}\, A_{ij}
        &=
        \int_{\cos\theta_{ij}}^{1}
        \frac{\dd \cos\theta_{iq}}{1-\cos\theta_{iq}}
        +
        \int_{\cos\theta_{ij}}^{1}
        \frac{\dd \cos\theta_{jq}}{1-\cos\theta_{jq}}
        \\
        &=
        2
        \le(
            \int_{0}^{\theta_{ij}}
            \frac{\dd \theta_{iq}}{\theta_{iq}}
            +
            \int_{0}^{\theta_{ij}}
            \frac{\dd \theta_{jq}}{\theta_{jq}}
        \ri)
        +
        \mathcal{O}(\theta_{ij}^2)
        \,,
    \end{align}
    where the first term of the right-hand side emerges after integrating over the azimuthal direction of \(\vec{q}\) relative to the axis defined by the motion of particle \(i\), and the second when integrating over the azimuthal angle relative to \(j\).
    %
    Notably, the phase space measure for \(\vec{q}\) of which vanishes when the direction of \(\vec{q}\) lies outside the cone defined by the momenta of \(i\) (in the first term) or \(j\) (in the second).
\end{lemma}

@@

    In the limit that \(q\) is collinear to particle 1, is it the first term in the parentheses of the right-hand side that matters.

@@

\begin{lemma}{Coherent Branching}{qcd-coherent-branching}
    Consider \(M\) outgoing particles, and an additional soft gluon, in the \vocab{strongly-angular-ordered limit}:%
    \footnote{
        Using the triangle equality, we could also write
        \(
            \theta_{12} \ll \theta_{(12)3} \ll \theta_{(123)4} \ll \cdots
            \,,
        \)
        where \(\theta_{(S)\,\,n}\) is the angle between \(\sum_{i\in S} \vec{k}_i\) and \(\vec{k}_n\).
    }
    \begin{align}
        \theta_{2} \ll \theta_{3} \ll \theta_{4}
        \ll
        \cdots
        \ll
        \theta_M
        \,,
    \end{align}
    where we define \(\theta_j = \theta_{1j}\) to be the angle of particle \(j\) relative to particle 1 for brevity.

@@

@@

\makeprob{The Cauchy Principle Value}{principle-value}{
    Provide a formal argument that
    \begin{align}
        \frac{\dd}{\dd x} \ln\le(\abs{x}\ri)
        =
        \text{PV}\le(\frac{1}{x} \ri)
        \,,
    \end{align}
    where the right-hand side is the \vocab{Cauchy Principle Value}, a distribution defined by
    \begin{align}
        \int_{-\infty}^{\infty} \dd x
        \,\,
        \text{PV}\le(\frac{1}{x} \ri)
        \,
        \phi(x)
        =
        \int_0^\infty
        \,
        \frac{\phi(x) - \phi(-x)}{x}
        \,
        \dd x
        \,.
    \end{align}
}
